---
title: "Covid Trash Models"
output: html_notebook
---

Models we are throwing away
###Florida Cases - Baseline Model One

To model Florida cases we start with a base model. We can see that the most favored model by BIC is an AR(1). So we begin by building that model. Our AR(1) has a phi of .975, quite close to the unit circle, which we expect to model strongly wandering behavior. 

This model has an AIC estimate of 14.01. 

In order to estimate an average ASE, we are running this model over segments of our data with 54 iterations. In each case training with at least seventy data points and predicting on twelve. Overall this produces an average ASE of 6,353,070. 

Below our ASE estimates we forecast short term and long term forecasts and both follow AR(1) behavior of data dampening towards our mean. 

```{r}
aic5.wge(newcases_fl$positiveIncrease, type = 'bic')


fl_ar_1 = est.ar.wge(newcases_fl$positiveIncrease, p = 1)

fl_ar_1$aic

trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(135-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(newcases_fl$positiveIncrease[i:(i+(trainingSize-1))],phi = fl_ar_1$phi, n.ahead = horizon, plot = FALSE)
  
  ASE = mean((newcases_fl$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for Model ARIMA(2,1,1) for Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE


short_fl_ar1 = fore.aruma.wge(newcases_us$positiveIncrease,phi = fl_ar_1$phi, n.ahead = 7, plot = TRUE)

long_us_ar1 = fore.aruma.wge(newcases_us$positiveIncrease,phi = fl_ar_1$phi, n.ahead = 90, plot = TRUE)



```


###US Model One - AR 1 

```{r}
aic5.wge(newcases_us$positiveIncrease)
aic5.wge(newcases_us$positiveIncrease, type = 'bic')


us_ar_1 = est.ar.wge(newcases_us$positiveIncrease, p = 1)
us_ar_1
us_ar_1$aic

trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(177-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(newcases_us$positiveIncrease[i:(i+(trainingSize-1))],phi = us_ar_1$phi, n.ahead = horizon, plot = FALSE)
  
  ASE = mean((newcases_us$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for Model ARIMA(2,1,1) for Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE


short_fl_ar1 = fore.aruma.wge(newcases_us$positiveIncrease,phi = us_ar_1$phi, n.ahead = 7, plot = TRUE)

long_us_ar1 = fore.aruma.wge(newcases_us$positiveIncrease,phi = us_ar_1$phi, n.ahead = 90, plot = TRUE)



```


####Florida ARIMA Model

```{r}

diff_fl_cases_1 = artrans.wge(newcases_fl$positiveIncrease, 1)
diff_fl_cases_2 = artrans.wge(diff_fl_cases_1, 1)

diff_fl_cases_2_7 = artrans.wge(diff_fl_cases_2, c(rep(0,6),1))

aic5.wge(diff_fl_cases_2_7)

est.arma.wge(diff_fl_cases_2_7)


model_arima = arima(newcases_fl$positiveIncrease, order = c(2,2,0), seasonal = list(order = c(7,0,0)))
model_arima



trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(135-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(newcases_fl$positiveIncrease[i:(i+(trainingSize-1))],phi = model_arima$coef[1:2], s = 7, d = 2, n.ahead = horizon, plot = FALSE)
  
  ASE = mean((newcases_fl$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for Model ARIMA(2,1,1) for Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
#25257410

short_fl_ar1 = fore.aruma.wge(newcases_fl$positiveIncrease,phi = model_arima$coef[1:2], s = 7, d = 2, n.ahead = 7, plot = TRUE)

long_fl_ar1 = fore.aruma.wge(newcases_fl$positiveIncrease,phi = model_arima$coef[1:2], s = 7, d = 2,  n.ahead = 90, plot = TRUE)



```


VAR/MLR with Correlated Errors Model  


```{r}
library(vars)
VARselect(newcases_fl_multi,lag.max = 10, type = "both")
#AIC of 25.76087

cases_fl_VAR = VAR(newcases_fl_multi,type = "trend",p = 4)
preds=predict(cases_fl_VAR,n.ahead=12)
plot(preds)



ASE = mean((newcases_fl_multi$positiveIncrease[123:135] - preds$fcst$positiveIncrease)^2)
ASE 
#35935880



```


