---
title: "Time_Series_Covid_Project"
author: "Reagan Meagher"
date: "7/13/2020"
output: html_document
---

**Initial Set Up and Data Read In:**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Global Options
remove(list = ls())

#Load Packages
library(tidyverse)
library(dplyr)
library(tseries)
library(ggplot2)
library(forecast)
library(tswge)
library(tseries)
library(orcutt)
library(mice)
```

```{r}
initial_data_fl <- read.csv(file="https://raw.githubusercontent.com/megnn/TimeSeries_Covid/master/covidtracking_FL_data.csv", header=TRUE)
initial_data_us <- read.csv(file="https://raw.githubusercontent.com/megnn/TimeSeries_Covid/master/covidtracking_US_data.csv", header=TRUE)

initial_data_fl = initial_data_fl[order(nrow(initial_data_fl):1),]
initial_data_us = initial_data_us[order(nrow(initial_data_us):1),]
```

```{r}
for (i in 1:nrow(initial_data_fl)) {
  n <- round((initial_data_fl$positiveIncrease / initial_data_fl$totalTestResultsIncrease) * 100, digits = 4)
  initial_data_fl$positive_percentage <- n
}

for (i in 1:nrow(initial_data_us)) {
  n <- round((initial_data_us$positiveIncrease / initial_data_us$totalTestResultsIncrease) * 100, digits = 4)
  initial_data_us$positive_percentage <- n
}

len_fl = dim(initial_data_fl)[1]
len_us = dim(initial_data_us)[1]
```

##**Daily Positive Cases Analysis:**

Set Up Data Sources:
```{r}
positiveCases_fl <- select(initial_data_fl, c("date", "positiveIncrease"))
positiveCases_us <- select(initial_data_us, c("date", "positiveIncrease"))

#Checking for NAs
md.pattern(positiveCases_fl)
#Drop NAs that are present
positiveCases_fl <- na.omit(positiveCases_fl)

#Checking for NAs
md.pattern(positiveCases_us)
#Drop NAs that are present
positiveCases_us <- na.omit(positiveCases_us)
```

###**Florida Daily Cases:**

Florida Data Plot:
```{r}
plot(x = seq(1,len_fl), y = positiveCases_fl$positiveIncrease, type = "l", main = 'Florida Covid-19 Daily Positive Cases', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Positive Daily Cases')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(positiveCases_fl$positiveIncrease)

aic5.wge(positiveCases_fl$positiveIncrease)
aic5.wge(positiveCases_fl$positiveIncrease, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for AR(1) Model:')
AR1_pc_fl = est.ar.wge(positiveCases_fl$positiveIncrease, p = 1)
AR1_pc_fl
print("Estimates for ARMA(2,1) Model:")
ARMA2_1_pc_fl = est.arma.wge(positiveCases_fl$positiveIncrease, p = 2, q = 1)
ARMA2_1_pc_fl
```

```{r}
print("AR(1) Two-Week Forecast")
f_AR1_pc_fl_s = fore.arma.wge(positiveCases_fl$positiveIncrease, phi=c(AR1_pc_fl$phi),n.ahead=14,lastn=FALSE)
print("ARMA(2,1) Two-Week Forecast")
f_ARMA2_1_pc_fl_s = fore.arma.wge(positiveCases_fl$positiveIncrease, phi=c(ARMA2_1_pc_fl$phi),theta=c(ARMA2_1_pc_fl$theta), n.ahead=14,lastn=FALSE)

print("AR(1) Two-Month Forecast")
f_AR1_pc_fl_l = fore.arma.wge(positiveCases_fl$positiveIncrease, phi=c(AR1_pc_fl$phi),n.ahead=60,lastn=FALSE)
print("ARMA(2,1) Two-Month Forecast")
f_ARMA2_1_pc_fl_l = fore.arma.wge(positiveCases_fl$positiveIncrease, phi=c(ARMA2_1_pc_fl$phi),theta=c(ARMA2_1_pc_fl$theta), n.ahead=60,lastn=FALSE)
```

###**National Daily Cases:**

United States Data Plot:
```{r}
plot(x = seq(1,len_us), y = positiveCases_us$positiveIncrease, type = "l", main = 'United States Covid-19 Daily Positive Cases', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Positive Daily Cases')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(positiveCases_us$positiveIncrease)

aic5.wge(positiveCases_us$positiveIncrease)
aic5.wge(positiveCases_us$positiveIncrease, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for AR(1) Model:')
AR1_pc_us = est.ar.wge(positiveCases_us$positiveIncrease, p = 1)
AR1_pc_us
print("Estimates for ARMA(1,2) Model:")
ARMA1_2_pc_us = est.arma.wge(positiveCases_us$positiveIncrease, p = 1, q = 2)
ARMA1_2_pc_us
```

Forecasts
```{r}
print("AR(1) Two-Week Forecast")
f_AR1_pc_us_s = fore.arma.wge(positiveCases_us$positiveIncrease, phi=c(AR1_pc_us$phi),n.ahead=14,lastn=FALSE)
print("ARMA(1,2) Two-Week Forecast")
f_ARMA2_1_pc_us_s = fore.arma.wge(positiveCases_us$positiveIncrease, phi=c(ARMA1_2_pc_us$phi),theta=c(ARMA1_2_pc_us$theta), n.ahead=14,lastn=FALSE)

print("AR(1) Two-Month Forecast")
f_AR1_pc_us_l = fore.arma.wge(positiveCases_us$positiveIncrease, phi=c(AR1_pc_us$phi),n.ahead=60,lastn=FALSE)
print("ARMA(1,2) Two-Month Forecast")
f_ARMA2_1_pc_us_l = fore.arma.wge(positiveCases_us$positiveIncrease, phi=c(ARMA1_2_pc_us$phi),theta=c(ARMA1_2_pc_us$theta), n.ahead=60,lastn=FALSE)
```

##**Daily Positive Percentage Analysis:**

Set Up Data Sources:
```{r}
positivePercentage_fl <- select(initial_data_fl, c("date", "positive_percentage"))
positivePercentage_us <- select(initial_data_us, c("date", "positive_percentage"))

#Checking for NAs
md.pattern(positivePercentage_fl)
#Drop NAs that are present
positivePercentage_fl <- na.omit(positivePercentage_fl)

#Checking for NAs
md.pattern(positivePercentage_us)
#Drop NAs that are present
positivePercentage_us <- na.omit(positivePercentage_us)
```

###**Florida Daily Positive Percentage:**

Florida Data Plot:
```{r}
len_fl_2 = dim(positivePercentage_fl)[1]
plot(x = seq(1,len_fl_2), y = positivePercentage_fl$positive_percentage, type = "l", main = 'Florida Covid-19 Daily Positive Percentage', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Daily Positive Percentage')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(positivePercentage_fl$positive_percentage)

aic5.wge(positivePercentage_fl$positive_percentage)
aic5.wge(positivePercentage_fl$positive_percentage, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for ARMA(3,1) Model:')
ARMA3_1_pp_fl = est.arma.wge(positivePercentage_fl$positive_percentage, p = 3, q = 1)
ARMA3_1_pp_fl
print("Estimates for ARMA(3,2) Model:")
ARMA3_2_pc_fl = est.arma.wge(positivePercentage_fl$positive_percentage, p = 3, q = 2)
ARMA3_2_pc_fl
```

```{r}
print("ARMA(3,1) Two-Week Forecast")
f_ARMA3_1_pp_fl_s = fore.arma.wge(positivePercentage_fl$positive_percentage, phi=c(ARMA3_1_pp_fl$phi), theta=c(ARMA3_1_pp_fl$theta),n.ahead=14,lastn=FALSE)
print("ARMA(2,1) Two-Week Forecast")
f_ARMA3_2_pc_fl_s = fore.arma.wge(positivePercentage_fl$positive_percentage, phi=c(ARMA3_2_pc_fl$phi),theta=c(ARMA3_2_pc_fl$theta), n.ahead=14,lastn=FALSE)

print("ARMA(3,1) Two-Month Forecast")
f_ARMA3_1_pp_fl_l = fore.arma.wge(positivePercentage_fl$positive_percentage, phi=c(ARMA3_1_pp_fl$phi),n.ahead=60,lastn=FALSE)
print("ARMA(2,1) Two-Month Forecast")
f_ARMA3_2_pc_fl_l = fore.arma.wge(positivePercentage_fl$positive_percentage, phi=c(ARMA3_2_pc_fl$phi),theta=c(ARMA3_2_pc_fl$theta), n.ahead=60,lastn=FALSE)
```

###**National Daily Positive Percentage:**

United States Data Plot:
```{r}
len_us_2 = dim(positivePercentage_us)[1]
plot(x = seq(1,len_us_2), y = positivePercentage_us$positive_percentage, type = "l", main = 'United States Covid-19 Daily Positive Percentage', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Daily Positive Percentage')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(positivePercentage_us$positive_percentage)

aic5.wge(positivePercentage_us$positive_percentage)
aic5.wge(positivePercentage_us$positive_percentage, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for ARMA(1,2) Model:')
ARMA1_2_pp_us = est.arma.wge(positivePercentage_us$positive_percentage, p = 1, q = 2)
ARMA1_2_pp_us
print("Estimates for ARMA(2,2) Model:")
ARMA2_2_pp_us = est.arma.wge(positivePercentage_us$positive_percentage, p = 2, q = 2)
ARMA2_2_pp_us
```

Forecasts
```{r}
print("ARMA(1,2) Two-Week Forecast")
f_ARMA1_2_pp_us_s = fore.arma.wge(positivePercentage_us$positive_percentage, phi=c(ARMA1_2_pp_us$phi),theta=c(ARMA1_2_pp_us$theta),n.ahead=14,lastn=FALSE)
print("ARMA(1,2) Two-Week Forecast")
f_ARMA2_2_pp_us_s = fore.arma.wge(positivePercentage_us$positive_percentage, phi=c(ARMA2_2_pp_us$phi),theta=c(ARMA2_2_pp_us$theta), n.ahead=14,lastn=FALSE)

print("ARMA(1,2) Two-Month Forecast")
f_ARMA1_2_pp_us_l = fore.arma.wge(positivePercentage_us$positive_percentage, phi=c(ARMA1_2_pp_us$phi),theta=c(ARMA1_2_pp_us$theta), n.ahead=60,lastn=FALSE)
print("ARMA(2,2) Two-Month Forecast")
f_ARMA2_2_pp_us_l = fore.arma.wge(positivePercentage_us$positive_percentage, phi=c(ARMA2_2_pp_us$phi),theta=c(ARMA2_2_pp_us$theta), n.ahead=60,lastn=FALSE)
```

##**Daily Deaths Analysis:**

Set Up Data Sources:
```{r}
deaths_fl <- select(initial_data_fl, c("date", "deathIncrease"))
deaths_us <- select(initial_data_us, c("date", "deathIncrease"))

#Checking for NAs
md.pattern(deaths_fl)
#Drop NAs that are present
deaths_fl <- na.omit(deaths_fl)

#Checking for NAs
md.pattern(deaths_us)
#Drop NAs that are present
deaths_us <- na.omit(deaths_us)
```

###**Florida Daily Cases:**

Florida Data Plot:
```{r}
plot(x = seq(1,len_fl), y = deaths_fl$deathIncrease, type = "l", main = 'Florida Covid-19 Daily Deaths', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Daily Deaths')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(deaths_fl$deathIncrease)

aic5.wge(deaths_fl$deathIncrease)
aic5.wge(deaths_fl$deathIncrease, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for ARMA(5,1) Model:')
ARMA5_1_d_fl = est.arma.wge(deaths_fl$deathIncrease, p = 5, q = 1)
ARMA5_1_d_fl
print("Estimates for ARMA(1,1) Model:")
ARMA1_1_d_fl = est.arma.wge(deaths_fl$deathIncrease, p = 1, q = 1)
ARMA1_1_d_fl
```

Forecasts
```{r}
print("ARMA(5,1) Two-Week Forecast")
f_ARMA5_1_d_fl_s = fore.arma.wge(deaths_fl$deathIncrease, phi=c(ARMA5_1_d_fl$phi),theta=c(ARMA5_1_d_fl$theta),n.ahead=14,lastn=FALSE)
print("ARMA(1,1) Two-Week Forecast")
f_ARMA1_1_d_fl_s = fore.arma.wge(deaths_fl$deathIncrease, phi=c(ARMA1_1_d_fl$phi),theta=c(ARMA1_1_d_fl$theta), n.ahead=14,lastn=FALSE)

print("ARMA(5,1) Two-Month Forecast")
f_ARMA5_1_d_fl_l = fore.arma.wge(deaths_fl$deathIncrease, phi=c(ARMA5_1_d_fl$phi),theta=c(ARMA5_1_d_fl$theta), n.ahead=60,lastn=FALSE)
print("ARMA(1,1) Two-Month Forecast")
f_ARMA1_1_d_fl_l = fore.arma.wge(deaths_fl$deathIncrease, phi=c(ARMA1_1_d_fl$phi),theta=c(ARMA1_1_d_fl$theta), n.ahead=60,lastn=FALSE)
```

###**National Daily Cases:**

United States Data Plot:
```{r}
plot(x = seq(1,len_us), y = deaths_us$deathIncrease, type = "l", main = 'United States Covid-19 Daily Deaths', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Daily Deaths')
```

Time Series Plots and Model Type Estimation:
```{r}
plotts.sample.wge(deaths_us$deathIncrease)

aic5.wge(deaths_us$deathIncrease)
aic5.wge(deaths_us$deathIncrease, type = 'bic')
```

Estimate Model Phi's/Theta's For Each Model:
```{r}
print('Estimates for ARMA(5,2) Model:')
ARMA5_2_d_us = est.arma.wge(deaths_us$deathIncrease, p = 5, q = 2)
ARMA5_2_d_us
print("Estimates for ARMA(4,2) Model:")
ARMA4_2_d_us = est.arma.wge(deaths_us$deathIncrease, p = 4, q = 2)
ARMA4_2_d_us
```

Forecasts
```{r}
print("ARMA(5,2) Two-Week Forecast")
f_ARMA5_2_d_us_s = fore.arma.wge(deaths_us$deathIncrease, phi=c(ARMA5_2_d_us$phi),theta=c(ARMA5_2_d_us$theta),n.ahead=14,lastn=FALSE)
print("ARMA(4,2) Two-Week Forecast")
f_ARMA4_2_d_us_s = fore.arma.wge(deaths_us$deathIncrease, phi=c(ARMA4_2_d_us$phi),theta=c(ARMA4_2_d_us$theta), n.ahead=14,lastn=FALSE)

print("ARMA(5,2) Two-Month Forecast")
f_ARMA5_2_d_us_l = fore.arma.wge(deaths_us$deathIncrease, phi=c(ARMA5_2_d_us$phi),theta=c(ARMA5_2_d_us$theta), n.ahead=60,lastn=FALSE)
print("ARMA(4,2) Two-Month Forecast")
f_ARMA4_2_d_us_l = fore.arma.wge(deaths_us$deathIncrease, phi=c(ARMA4_2_d_us$phi),theta=c(ARMA4_2_d_us$theta), n.ahead=60,lastn=FALSE)
```

