---
title: "Time Series Covid Project"
author: "Reagan Meagher and Megan Riley"
output:
  pdf_document: default
  html_document: default
---

[site]: https://covidtracking.com/ "The Covid Tracking Project"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Global Options
remove(list = ls())
options(scipen = 999)

#Load Packages
library(tidyverse)
library(dplyr)
library(tseries)
library(ggplot2)
library(forecast)
library(tswge)
library(tseries)
library(orcutt)
library(mice)
library(nnfor)
```

#*Modeling Covid19 Time Series* 


Covid19 is a worldwide pandemic that will likely define 2020. In the United States, currently over four million people have been infected and over 150,000 have died as a result of Covid19. As the pandemic continues, limiting infections, serious harm, and death is a primary concern for all involved. 

As this is a novel illness, we know relatively little, but understanding how Covid19 is spreading and judging the severity of an outbreak can be approximated with the data we have available. In this report we aim to build effective time series models to forecast future Covid19 cases using the techniques we have learned from this Time Series course. 
##*Goal One:  Data Collection *

The data source we are using is sourced from [The Covid Tracking Project][site]. 



```{r}
initial_data_fl <- read.csv(file="https://raw.githubusercontent.com/megnn/TimeSeries_Covid/master/covidtracking_FL_data.csv", header=TRUE)
initial_data_us <- read.csv(file="https://raw.githubusercontent.com/megnn/TimeSeries_Covid/master/covidtracking_US_data.csv", header=TRUE)

initial_data_fl = initial_data_fl[order(nrow(initial_data_fl):1),]
initial_data_us = initial_data_us[order(nrow(initial_data_us):1),]

len_fl = dim(initial_data_fl)[1]
len_us = dim(initial_data_us)[1]

```


Below we plot the realizations of daily new cases from both Florida and the United States as a whole. 

```{r}
plot(x = seq(1,len_fl), y = initial_data_fl$positiveIncrease, type = "l", ylab = "Count of Cases", xlab = "Time", main ="Count of Daily Covid19 Cases - Florida")
plot(x = seq(1,len_us), y = initial_data_us$positiveIncrease, type = "l", ylab = "Count of Cases", xlab = "Time", main ="Count of Daily Covid19 Cases - United States")
```

##*Positive Percentage*

Positive Percentage is a statistic that calculates daily positive tests as a percentage of daily overall tests returned. We calculated this column and added it to our data below followed by some visual exploration of the statistic itself. 

Overall we see a clear instance of high and often 100% positive test rates early on in the first days and weeks of the pandemic spread. We understand this as a result of the fact that Covid19 spread fast and we had more community spread than anticipated early on without the testing available. It is abundantly clear that when we have extremely high percent positive rates near 100% we can expect true positive case numbers at the time to be under represented. But without better epidemlogical understanding we can't make judgement calls on true case numbers when percent positives rise from 5% to 10% as we see begin to happen somewhat in recent days in Florida. 


```{r}
for (i in 1:nrow(initial_data_fl)) {
  n <- round((initial_data_fl$positiveIncrease / initial_data_fl$totalTestResultsIncrease) * 100, digits = 4)
  initial_data_fl$positive_percentage <- n
}

for (i in 1:nrow(initial_data_us)) {
  n <- round((initial_data_us$positiveIncrease / initial_data_us$totalTestResultsIncrease) * 100, digits = 4)
  initial_data_us$positive_percentage <- n
}


#Percent Positive Exploration
plot(x = seq(1:len_fl), y = initial_data_fl$positive_percentage, type = "l", main = "Florida Positive Percentage over time", ylab = "Positive Percentage", xlab = "Time")



plot(x = seq(1:len_us), y = initial_data_us$positive_percentage, type = "l", main = "US Positive Percentage over time", ylab = "Positive Percentage", xlab = "Time")



```



Positive Percentage as a metric is a measure of two main things, how many tests are we administering and how many positives are we receiving. If tests are skyrocketing while positive cases are increasing, we would see a stable or even diminishing line which could indicate not a pandemic under control but simply better testing resources but could be interpreted as a pandemic managed. 

Keeping tests increasing to continue to keep percent positives level is a good indication we have leveled up our resources to continue to diagnose the pandemic at the same level, but if we need to scale up our testing to keep the same positive percentage, there is more covid spread. 

However, an increasing positive percentage is a good indicator that our testing resources may not be up to actually up to tracking the current stage of the pandemic. 


```{r}
plot(x = seq(1:len_us), y = initial_data_us$totalTestResultsIncrease, type = "l")
plot(x = seq(1:len_fl), y = initial_data_fl$totalTestResultsIncrease, type = "l")

```

####Data Preperation

In order to model new case numbers by day we set up dataframes with only our date and positive increase amount per day.

```{r}
newcases_fl <- dplyr::select(initial_data_fl, c("date", "positiveIncrease"))
newcases_us <- dplyr::select(initial_data_us, c("date", "positiveIncrease"))
```



####Checking for NAs

We can see with the missing value analysis below that we have no NAs present in our new case data. 

```{r NA eval}
#Checking for NAs
md.pattern(newcases_fl)
# No NAs present


#Checking for NAs
md.pattern(newcases_us)
#No NAs present

```


### **Florida Daily Cases:**


```{r results='hide'}
#no text output, just the plots
x = plotts.sample.wge(newcases_fl$positiveIncrease)

```


###**US Daily Cases:**



```{r results='hide'}
#no text output
x = plotts.sample.wge(newcases_us$positiveIncrease)

dev.off()
plot(x = seq(1,len_us), y = newcases_us$positiveIncrease, type = "l", main = 'United States Covid-19 Daily Positive Cases', xlab = 'Time (Days From Start of Pandemic)', ylab = 'Positive Daily Cases')
 
```




##*Goal Two: Univariate Analysis*

###Model Building for Cases in Florida


A. Stationarity vs Non-Stationarity

Overall we see slowly dampening ACFs, combined with a strong aperiodic frequency at zero in our spectral density. These measures alone with a recently quickly rising case count in recent days gives us strong evidence that our data is non-stationary. Given Covid19 spread, it is likely we see continued rising behavior in the short term, some return to lower numbers in the coming months but more uncertainty as new spikes could arise, and in the longest term of years on, we expect new cases to diminish to zero once the pandemic has ended spread. 

```{r results='hide'}
x = plotts.sample.wge(newcases_fl$positiveIncrease, arlimits = TRUE)

```

B. Non Stationary Modeling 

We did not do any differencing of our data set to account for this non-stationarity. Going into this project we knew that because of the failure to contain the Covid-19 outbreak we would see large spikes of cases in recent time periods compared to distant time periods. We feel that this is an important aspect of our data that we want to portray in our models because we can see empirically in Florida and the United States as a whole that both individual behavior and political policy continue to trend towards further outbreak and rapid, almost exponential daily case growth. While some states with compliant individual behavior and strong political Covid-19 policies have shown “completed” Covid-19 curves, where daily case count begins to trend downwards towards zero, Florida is the opposite. Therefore, since we empirically expect the trend of non-stationarity to continue, we want that represented in our models. This is a fundamental assumption that our models are built on.



```{r}

diff_fl = artrans.wge(newcases_fl$positiveIncrease, 1)

```

C. Model IDing of stationary models

For ARIMA models we identify the stationary components below. 

```{r}
#Model Differenced Data
aic5.wge(diff_fl)
aic5.wge(diff_fl, type = "bic") #2,2 produced


#Modeling original data if stationary
aic5.wge(newcases_fl$positiveIncrease)

aic5.wge(newcases_fl$positiveIncrease, type = 'bic')

```


D. Model Building 


#### Florida Cases - ARIMA Model


```{r}

fl_arima = est.arma.wge(diff_fl, p = 2, q = 2)

fl_arima$aic

trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(135-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(newcases_fl$positiveIncrease[i:(i+(trainingSize-1))],phi = fl_arima$phi, theta = fl_arima$theta, d = 1, n.ahead = horizon, plot = FALSE)
  
  ASE = mean((newcases_fl$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for Model ARIMA(2,1,2) for Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
##



short_fl_arima = fore.aruma.wge(newcases_fl$positiveIncrease,phi = fl_arima$phi,theta = fl_arima$theta, d=1, n.ahead = 7, plot = TRUE)

long_fl_arima = fore.aruma.wge(newcases_fl$positiveIncrease,phi = fl_arima$phi,theta = fl_arima$theta, d=1 ,n.ahead = 90, plot = TRUE)

final_pred = fore.aruma.wge(newcases_fl$positiveIncrease[1:123],phi = fl_arima$phi,theta = fl_arima$theta, d=1, n.ahead = 12, plot = TRUE)
final_pred_df = data.frame(t = seq(124:135), final_pred$f)

plot(newcases_fl$positiveIncrease, type = "l", ylab = "Count of New Cases", xlab = "Time", main = "Florida ARIMA(2,1,2) Model Final 12 Predictions", xlim = c(1,140))
lines(ts(final_pred$f, start = 124, end = 135),  col = "blue")

final_ASE = mean((newcases_fl$positiveIncrease[124:135] - final_pred$f)^2)
final_ASE



```





####MLP Model for Florida Cases

```{r}



trainingSize = 70
horizon = 12
ASEHolder = numeric()


for( i in 1:(135-(trainingSize + horizon) + 1))
{
  mlp.fit = mlp(ts(newcases_fl$positiveIncrease[1:trainingSize+i]), hd = 5, comb = "median")
  forecasts = forecast(mlp.fit,h = horizon)
  
  ASE = mean((newcases_fl$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -forecasts$mean)^2)
  
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLP Model Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE

plot(forecasts)

#Final Forecasts with data known
mlp.fit_fl_final = mlp(ts(newcases_fl$positiveIncrease[1:123]), hd = 5, comb = "median")
forecasts_fl_mlp = forecast(mlp.fit,h = 12)

final12_ase = mean((newcases_fl$positiveIncrease[124:135] -forecasts_fl_mlp$mean)^2)
final12_ase

all_f = c(rep(1,4), forecasts$fitted, forecasts$mean)
plot(newcases_fl$positiveIncrease, type = "l", ylab = "New Daily Cases", xlab = "Day", main = "")
lines(all_f, col = "blue")


#Future Forecasts

mlp.fit_fl_future =mlp(ts(newcases_fl$positiveIncrease), hd = 5, comb = "median")

short_fl_mlp = forecast(mlp.fit_fl_future,h = 7)
long_fl_mlp = forecast(mlp.fit_fl_future,h = 90)

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,145), main = "Florida Short Term MLP Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_fl_mlp$mean, col = "blue")

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,235), ylim = c(0,23000), main = "Florida Long Term MLP Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_fl_mlp$mean, col = "blue")




```



####Florida Cases Ensemble Model

```{r}
#ASE fits for ensemble
mlp.fit_fl_final = mlp(ts(newcases_fl$positiveIncrease[1:123]), hd = 5, comb = "median")
forecasts_fl_mlp = forecast(mlp.fit_fl_final, h = 12)

forecasts_fl_arima = fore.aruma.wge(newcases_fl$positiveIncrease[i:(i+(trainingSize-1))],phi = fl_arima$phi, theta = fl_arima$theta, d=1, n.ahead = 12, plot = FALSE)

ensemble_fl_fore = (forecasts_fl_mlp$mean + final_pred_df$f) / 2

ensemble_ASE = mean((newcases_fl$positiveIncrease[124:135] - ensemble_fl_fore)^2)
ensemble_ASE
#8.4 Mill

plot(newcases_fl$positiveIncrease, type = "l", ylab = "Count of New Cases", xlab = "Time", main = "Florida Ensemble Model Final 12 Predictions")
lines(ensemble_fl_fore, col = "blue")


#future forecasting
 
short_fl_ensemble = (short_fl_mlp$mean + short_fl_arima$f)/2
long_fl_ensemble = (long_fl_mlp$mean + long_fl_arima$f)/2



plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,145), main = "Florida Short Term Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_fl_ensemble, col = "blue")

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,235), ylim = c(0,23000), main = "Florida Long Term Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_fl_ensemble, col = "blue")



```



Comparing and Assessing Models

```{r}

```

###Model Building for Cases United States


A. Stationarity vs Non-Stationarity

Overall we see slowly dampening ACFs, combined with a strong aperiodic frequency at zero in our spectral density. These measures alone with a recently quickly rising case count in recent days gives us strong evidence that our data is non-stationary. Given Covid19 spread, it is likely we see continued rising behavior in the short term, some return to lower numbers in the coming months but more uncertainty as new spikes could arise, and in the longest term of years on, we expect new cases to diminish to zero once the pandemic has ended spread. 

```{r results='hide'}

#no text output
x = plotts.sample.wge(newcases_us$positiveIncrease)

```

B. Non-Stationary Modeling

We did not do any differencing of our data set to account for this non-stationarity. Going into this project we knew that because of the failure to contain the Covid-19 outbreak we would see large spikes of cases in recent time periods compared to distant time periods. We feel that this is an important aspect of our data that we want to portray in our models because we can see empirically in Florida and the United States as a whole that both individual behavior and political policy continue to trend towards further outbreak and rapid, almost exponential daily case growth. While some states with compliant individual behavior and strong political Covid-19 policies have shown “completed” Covid-19 curves, where daily case count begins to trend downwards towards zero, Florida is the opposite. Therefore, since we empirically expect the trend of non-stationarity to continue, we want that represented in our models. This is a fundamental assumption that our models are built on.

```{r}

diff_us = artrans.wge(newcases_us$positiveIncrease, 1)

```

C. Model IDing of stationary models 

For ARIMA models we identify the stationary components below. 

```{r}
#modeling as non-stationary
#0,5 maxes out
aic5.wge(diff_us, p = 3:10)
aic5.wge(diff_us, type = 'bic',p = 3:10)


#Modeling as stationary
aic5.wge(newcases_us$positiveIncrease)
aic5.wge(newcases_us$positiveIncrease, type = 'bic')

```

D. Model Building 


####US Cases ARIMA

```{r}


us_arima = est.arma.wge(diff_us, p = 6, q = 2)
us_arima$aic

trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(177-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(newcases_us$positiveIncrease[i:(i+(trainingSize-1))],phi = us_arima$phi, theta = us_arima$theta, d = 1, n.ahead = horizon, plot = FALSE)
  
  ASE = mean((newcases_us$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for Model ARIMA(6,1,2) for US Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
#26724396


short_us_arima = fore.aruma.wge(newcases_us$positiveIncrease,phi = us_arima$phi,theta = us_arima$theta, d = 1, n.ahead = 7, plot = TRUE)

long_us_arima = fore.aruma.wge(newcases_us$positiveIncrease,phi = us_arima$phi,theta = us_arima$theta, d = 1, n.ahead = 90, plot = TRUE)



final_pred = fore.aruma.wge(newcases_us$positiveIncrease[1:165],phi = us_arima$phi,theta = us_arima$theta, d = 1, n.ahead = 12, plot = TRUE)

final_12_ase = mean((newcases_us$positiveIncrease[166:177] - final_pred$f)^2)
final_12_ase

plot(newcases_us$positiveIncrease, type = "l", ylab = "Count of New Cases", xlab = "Time", main = "United States ARIMA(6,1,2) Model Final 12 Predictions")
lines(ts(final_pred$f, start = 166, end = 177),  col = "blue")



```



####US Cases MLP 
```{r}
trainingSize = 70
horizon = 12
ASEHolder = numeric()


for( i in 1:(177-(trainingSize + horizon) + 1))
{
  mlp.fit = mlp(ts(newcases_us$positiveIncrease[1:trainingSize+i]), hd = 5, reps = 20, lags = c(1,3,4), comb = "median")
  forecasts = forecast(mlp.fit,h = horizon)
  
  ASE = mean((newcases_us$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -forecasts$mean)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLP Model United States Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
#228 mill

plot(forecasts)

#Actual Forecasting on last segment of data

mlp.fit = mlp(ts(newcases_us$positiveIncrease[1:165]), hd = 5, comb = "median")
forecasts_us_mlp = forecast(mlp.fit,h = 12)
  
ASE = mean((newcases_us$positiveIncrease[166:177] -forecasts_us_mlp$mean)^2)
ASE
#53,843,551
plot(newcases_us$positiveIncrease, type = "l")
lines(forecasts_us_mlp$fitted, col = "blue")


all_f = c(rep(1,4),forecasts_us_mlp$fitted, forecasts_us_mlp$mean)
plot(newcases_us$positiveIncrease, type = "l", ylab = "Count of Cases", xlab = "Time", main = "US MLP Cases Model")
lines(all_f, col = "blue")

#Future Predictions

mlp.fit_us_future =mlp(ts(newcases_us$positiveIncrease), hd = 5, comb = "median")
short_us_mlp = forecast(mlp.fit_us_future,h = 7)

long_us_mlp = forecast(mlp.fit_us_future,h = 90)

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,187), main = "United States Short Term MLP Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_us_mlp$mean, col = "blue")

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,277), main = "United States Long Term MLP Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_us_mlp$mean, col = "blue")



```


####US Ensemble
```{r}
#ASE fits for ensemble
#mlp.fit_us_final = mlp(ts(newcases_us$positiveIncrease[1:165]), hd = 5, comb = "median")
#forecasts_us_mlp = forecast(mlp.fit_us_final,h = 12)


forecasts_arima_us = fore.aruma.wge(newcases_us$positiveIncrease[1:165],phi = us_arima$phi,theta = us_arima$theta, d = 1, n.ahead = 12, plot = TRUE)

ensemble_fore = (forecasts_us_mlp$mean + forecasts_arima_us$f) / 2

ensemble_ASE = mean((newcases_us$positiveIncrease[166:177] -ensemble_fore)^2)
ensemble_ASE
#

plot(newcases_us$positiveIncrease, type = "l", ylab = "Count of New Cases", xlab = "Time")
lines(ensemble_fore, col = "blue")

#Forecasting ahead
short_ensemble_us = (short_us_mlp$mean +short_us_arima$f)/2

long_ensemble_us = (long_us_mlp$mean +long_us_arima$f)/2


plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,187), main = "United States Short Term Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_ensemble_us, col = "blue")

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,277), main = "United States Long Term Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_ensemble_us, col = "blue")




```




Model Building 

##*Goal Three: Multivariate Analysis*




```{r}

newcases_fl_multi =  initial_data_fl %>% dplyr::select(positiveIncrease, totalTestResultsIncrease, hospitalizedIncrease, deathIncrease)

#Forecast beyond data for Florida

#Forecast future variables
fit.mlp.1 = mlp(ts(newcases_fl_multi$totalTestResultsIncrease),reps = 20, comb = "median")
plot(fit.mlp.1)
fore.mlp.1 = forecast(fit.mlp.1, h = 100)
plot(fore.mlp.1)

fit.mlp.2 = mlp(ts(newcases_fl_multi$hospitalizedIncrease),reps = 20, comb = "median")
plot(fit.mlp.2)
fore.mlp.2 = forecast(fit.mlp.2, h = 100)
plot(fore.mlp.2)

fit.mlp.3 = mlp(ts(newcases_fl_multi$deathIncrease),reps = 20, comb = "median")
plot(fit.mlp.3)
fore.mlp.3 = forecast(fit.mlp.3, h = 100)
plot(fore.mlp.3)



#package them up in data frame.
newvar_fore_fl = data.frame(totalTestResultsIncrease = ts(c(newcases_fl_multi$totalTestResultsIncrease,fore.mlp.1$mean)), hospitalizedIncrease = ts(c(newcases_fl_multi$hospitalizedIncrease,fore.mlp.2$mean)), deathIncrease = ts(c(newcases_fl_multi$deathIncrease,fore.mlp.3$mean)))

#Data has 100 instances beyond current data
dim(newvar_fore_fl)



```

###Multivariate Model Building for Florida Cases



A. Stationarity vs Non-Stationarity

Overall we see slowly dampening ACFs, combined with a strong aperiodic frequency at zero in our spectral density. These measures alone with a recently quickly rising case count in recent days gives us strong evidence that our data is non-stationary. Given Covid19 spread, it is likely we see continued rising behavior in the short term, some return to lower numbers in the coming months but more uncertainty as new spikes could arise, and in the longest term of years on, we expect new cases to diminish to zero once the pandemic has ended spread. 

```{r results='hide'}
x = plotts.sample.wge(newcases_fl$positiveIncrease, arlimits = TRUE)

```

B. Non-Stationary Modeling

We did not do any differencing of our data set to account for this non-stationarity. Going into this project we knew that because of the failure to contain the Covid-19 outbreak we would see large spikes of cases in recent time periods compared to distant time periods. We feel that this is an important aspect of our data that we want to portray in our models because we can see empirically in Florida and the United States as a whole that both individual behavior and political policy continue to trend towards further outbreak and rapid, almost exponential daily case growth. While some states with compliant individual behavior and strong political Covid-19 policies have shown “completed” Covid-19 curves, where daily case count begins to trend downwards towards zero, Florida is the opposite. Therefore, since we empirically expect the trend of non-stationarity to continue, we want that represented in our models. This is a fundamental assumption that our models are built on.

C. Model ID 

In multivariate modeling, our identification of models occured specifically for each model and can be found at the beginning of those sections in particular. 

D. Model Building 

####Florida MLR Model
```{r}
fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_fl_multi)
summary(fit)
aic5.wge(fit$residuals)#picks 1,1

est1 = est.arma.wge(fit$residuals, p = 1, q = 1)


est_tests = mean(tail(newcases_fl_multi$totalTestResultsIncrease))
est_hospital= mean(tail(newcases_fl_multi$hospitalizedIncrease))

for( i in 1:(135-(trainingSize + horizon) ))
{
  fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_fl_multi[1:trainingSize+i,])
  newdata = data.frame(totalTestResultsIncrease = rep(est_tests,horizon), hospitalizedIncrease = rep(est_hospital,horizon))
  
  preds = predict(fit, newdata = newdata)
  forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = horizon, plot = FALSE)
  
  final_pred  = preds + forecasts$f

  
  ASE = mean((newcases_fl_multi$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -final_pred)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLR Model Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE



fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_fl_multi)
summary(fit)
newdata = data.frame(totalTestResultsIncrease = rep(est_tests,12), hospitalizedIncrease = rep(est_hospital,12))

preds = predict(fit, newdata = newdata)
aic5.wge(fit$residuals)#picks 1,1

est1 = est.arma.wge(fit$residuals, p = 1, q = 1)

forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 12)
FinalPredictions_fl_MLR  = preds + forecasts$f


plot(newcases_fl$positiveIncrease, type = "l", main ="Florida Multivariate MLR with Correlated Residuals Model", xlab = "Time", ylab = "Count of Cases")
lines(ts(FinalPredictions_fl_MLR, start = 124), col = "blue")


ASE = mean((newcases_fl_multi$positiveIncrease[124:135] - FinalPredictions_fl_MLR)^2)
ASE

  

#Forecasting Ahead

shortdata = data.frame(totalTestResultsIncrease = rep(est_tests,7), hospitalizedIncrease = rep(est_hospital,7))
longdata =data.frame(totalTestResultsIncrease = rep(est_tests,90), hospitalizedIncrease = rep(est_hospital,90))




fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_fl_multi)
#short
preds = predict(fit, newdata = shortdata)
forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 7)
short_fl_mlr_m  = preds + forecasts$f

#long
preds = predict(fit, newdata = longdata)
forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 90)
long_fl_mlr_m  = preds + forecasts$f


plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,145), main = "Florida Short Term MLR Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ts(short_fl_mlr_m, start = 136), col = "blue")

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,235), main = "Florida States Long Term MLR Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ts(long_fl_mlr_m,start = 136), col = "blue")


```



####Florida Multivariate MLP Cases Model

```{r}

newcases_fl_multi =  initial_data_fl %>% dplyr::select(positiveIncrease, totalTestResultsIncrease, hospitalizedIncrease, deathIncrease)
newcases_fl_var = cbind(ts(newcases_fl_multi$totalTestResultsIncrease),ts(newcases_fl_multi$hospitalizedIncrease),ts(newcases_fl_multi$deathIncrease))

trainingSize = 70
horizon = 12
ASEHolder = numeric()

#Out of bounds if it goes for 54 runs, this ASE will be slightly less wide than the others. But the windowed portion means its average ASE is still good
for( i in 1:(135-(trainingSize + horizon) ))
{
  mlp.fit = mlp(ts(newcases_fl_multi$positiveIncrease[1:trainingSize+i]), hd = 5, comb = "median", xreg = newcases_fl_var[1:trainingSize+i,])
  forecasts = forecast(mlp.fit,h = horizon, xreg = newcases_fl_var[1:(trainingSize + i + 12),])
  
  ASE = mean((newcases_fl_multi$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -forecasts$mean)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLP Model Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
#18757436 - 18 mill

plot(forecasts)

#Final Forecasts with data known
mlp.fit = mlp(ts(newcases_fl_multi$positiveIncrease[1:123]), hd = 5, comb = "median", xreg =newcases_fl_var[1:123,] )
forecasts = forecast(mlp.fit,h = 12, xreg = newcases_fl_var[1:135,])
fl_multi_mlp_fore = forecasts$mean

all_f = c(forecasts$fitted, forecasts$mean)
plot(newcases_fl_multi$positiveIncrease, type = "l", main = "Florida Multivariate MLP Model with Fits and Final 12 Predictions", xlab = "Time", ylab = "Count of Cases")
lines(all_f, col = "blue")

#Forecast beyond data
mlp.fit = mlp(ts(newcases_fl_multi$positiveIncrease), hd = 5, comb = "median", xreg = newvar_fore_fl[1:135,])
short_fl_mlp_m = forecast(mlp.fit,h = 7, xreg = newvar_fore_fl[1:145,])
long_fl_mlp_m = forecast(mlp.fit,h = 90, xreg = newvar_fore_fl[1:225,])

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,145),ylim = c(0,21000), main = "Florida Short Term MLP Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_fl_mlp_m$mean, col = "blue")

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,245),ylim = c(0,69000), main = "Florida Long Term MLP Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_fl_mlp_m$mean, col = "blue")


final_12_ase = mean((newcases_fl_multi$positiveIncrease[124:135] -fl_multi_mlp_fore)^2)
final_12_ase


```


####Florida Multivariate Ensemble Model

```{r}

ensemble_fore = (fl_multi_mlp_fore + FinalPredictions_fl_MLR)/2

plot(newcases_fl_multi$positiveIncrease, type = "l", main = "Florida Multivariate Ensemble Model with Final 12 Predictions", xlab = "Time", ylab = "Count of Cases")
lines(ensemble_fore, col = "blue")

ASE_fl_multi = mean((newcases_fl_multi$positiveIncrease[124:135] -ensemble_fore)^2)
ASE_fl_multi
#ASE of 8,427,522

#future

#long_fl_mlp_m
#short_fl_mlp_m


ensemble_fl_fore_short = ( short_fl_mlp_m$mean+ short_fl_mlr_m)/2
ensemble_fl_fore_long = (long_fl_mlp_m$mean + long_fl_mlr_m)/2

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,145),ylim = c(0,21000), main = "Florida Short Term Multivariate Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ensemble_fl_fore_short, col = "blue")

plot(newcases_fl$positiveIncrease, type = "l", xlim = c(1,235), ylim = c(0,40000),main = "Florida Long Term Multivariate Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ensemble_fl_fore_long, col = "blue")


```


###Multivariate US Models

A. Stationarity vs Non-Stationarity

Overall we see slowly dampening ACFs, combined with a strong aperiodic frequency at zero in our spectral density. These measures alone with a recently quickly rising case count in recent days gives us strong evidence that our data is non-stationary. Given Covid19 spread, it is likely we see continued rising behavior in the short term, some return to lower numbers in the coming months but more uncertainty as new spikes could arise, and in the longest term of years on, we expect new cases to diminish to zero once the pandemic has ended spread. 

```{r results='hide'}

#no text output
x = plotts.sample.wge(newcases_us$positiveIncrease)

```

B. Non-Stationary Modeling

We did not do any differencing of our data set to account for this non-stationarity. Going into this project we knew that because of the failure to contain the Covid-19 outbreak we would see large spikes of cases in recent time periods compared to distant time periods. We feel that this is an important aspect of our data that we want to portray in our models because we can see empirically in Florida and the United States as a whole that both individual behavior and political policy continue to trend towards further outbreak and rapid, almost exponential daily case growth. While some states with compliant individual behavior and strong political Covid-19 policies have shown “completed” Covid-19 curves, where daily case count begins to trend downwards towards zero, Florida is the opposite. Therefore, since we empirically expect the trend of non-stationarity to continue, we want that represented in our models. This is a fundamental assumption that our models are built on.


C. Model IDing of stationary models 

In multivariate modeling, our identification of models occured specifically for each model and can be found at the beginning of those sections in particular. 



D. Model Building 

In order to forecast multivariate models of new case numbers, we will first fit some new variables for future MLP type models. 

```{r}
newcases_us_multi =  initial_data_us %>% dplyr::select(positiveIncrease, totalTestResultsIncrease, hospitalizedIncrease, deathIncrease)

#Forecast Future


#Forecast future variables
fit.mlp.1 = mlp(ts(newcases_us_multi$totalTestResultsIncrease),reps = 20, comb = "median")
plot(fit.mlp.1)
fore.mlp.1 = forecast(fit.mlp.1, h = 100)
plot(fore.mlp.1)

fit.mlp.2 = mlp(ts(newcases_us_multi$hospitalizedIncrease),reps = 20, comb = "median")
plot(fit.mlp.2)
fore.mlp.2 = forecast(fit.mlp.2, h = 100)
plot(fore.mlp.2)

fit.mlp.3 = mlp(ts(newcases_us_multi$deathIncrease),reps = 20, comb = "median")
plot(fit.mlp.3)
fore.mlp.3 = forecast(fit.mlp.3, h = 100)
plot(fore.mlp.3)



#package them up in data frame.
newvar_fore_us = data.frame(totalTestResultsIncrease = ts(c(newcases_us_multi$totalTestResultsIncrease,fore.mlp.1$mean)), hospitalizedIncrease = ts(c(newcases_us_multi$hospitalizedIncrease,fore.mlp.2$mean)), deathIncrease = ts(c(newcases_us_multi$deathIncrease,fore.mlp.3$mean)))

dim(newvar_fore_us)


```


####US MLR with Correlated Errors Model  
```{r}
fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_us_multi[1:165,])
summary(fit)
est_tests = mean(tail(newcases_us_multi$totalTestResultsIncrease))
est_hospital= mean(tail(newcases_us_multi$hospitalizedIncrease))
newdata = data.frame(totalTestResultsIncrease = rep(est_tests,12), hospitalizedIncrease = rep(est_hospital,12))

preds = predict(fit, newdata = newdata)



aic5.wge(fit$residuals)#picks 3,2 with full data

est1 = est.arma.wge(fit$residuals, p = 3, q = 2)

for( i in 1:(177-(trainingSize + horizon) ))
{
  fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_us_multi[1:trainingSize+i,])
  newdata = data.frame(totalTestResultsIncrease = rep(est_tests,horizon), hospitalizedIncrease = rep(est_hospital,horizon))
  
  preds = predict(fit, newdata = newdata)
  forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = horizon, plot = FALSE)
  
  final_pred  = preds + forecasts$f

  
  ASE = mean((newcases_us_multi$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -final_pred)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLR Model United States Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE


fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_us_multi[1:165,])
preds = predict(fit, newdata = newdata)

forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 12)
FinalPredictions_us_MLR  = preds + forecasts$f

plot(newcases_us$positiveIncrease, type = "l", main ="US Multivariate MLR with Correlated Residuals Model", xlab = "Time", ylab = "Count of Cases")
lines(ts(FinalPredictions_us_MLR, start = 166), col = "blue")


ASE = mean((newcases_us_multi$positiveIncrease[166:177] - FinalPredictions_us_MLR)^2)
ASE
#

#Forecasting Ahead

shortdata = data.frame(totalTestResultsIncrease = rep(est_tests,7), hospitalizedIncrease = rep(est_hospital,7))
longdata =data.frame(totalTestResultsIncrease = rep(est_tests,90), hospitalizedIncrease = rep(est_hospital,90))


fit = lm(positiveIncrease~totalTestResultsIncrease + hospitalizedIncrease, data = newcases_us_multi)
#short
preds = predict(fit, newdata = shortdata)
forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 7)
short_us_mlr_m  = preds + forecasts$f

#long
preds = predict(fit, newdata = longdata)
forecasts = fore.arma.wge(fit$residuals,phi = est1$phi,theta = est1$theta, lastn = FALSE,n.ahead = 90)
long_us_mlr_m  = preds + forecasts$f


plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,187), main = "United States Short Term MLR Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ts(short_us_mlr_m, start = 178), col = "blue")

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,277), main = "United States Long Term MLR Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ts(long_us_mlr_m,start = 178), col = "blue")

```

####US MLP/RNN Model
```{r}



trainingSize = 70
horizon = 12
ASEHolder = numeric()

for( i in 1:(177-(trainingSize + horizon) + 1))
{
  mlp.fit = mlp(ts(newcases_us_multi$positiveIncrease[1:trainingSize+i]), hd = 5, comb = "median", xreg = newvar_fore_us[1:trainingSize+i,])
  forecasts = forecast(mlp.fit,h = horizon, xreg = newvar_fore_us[1:(trainingSize + i + 13),])
  
  ASE = mean((newcases_us_multi$positiveIncrease[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] -forecasts$mean)^2)
  ASEHolder[i] = ASE
  
}

ASEHolder
#Distribution of ASEs on Two Week Periods
hist(ASEHolder, xlab = "ASE of model at a given Training Set",  main = "ASE Distribution for MLP Model Florida Data")

#Mean ASE 
WindowedASE = mean(ASEHolder)
WindowedASE
#97494363

plot(forecasts)

#Final Forecasts with data known
mlp.fit = mlp(ts(newcases_us_multi$positiveIncrease[1:177]), hd = 5, comb = "median", xreg = newvar_fore_us[1:177,])
forecasts_us_mlp = forecast(mlp.fit,h = 12, xreg = newvar_fore_us[1:190,])

all_f = c(rep(1,4),forecasts_us_mlp$fitted, forecasts_us_mlp$mean)
plot(newcases_us_multi$positiveIncrease, type = "l")
lines(all_f, col = "blue")



#final 12 forecasts

mlp.fit = mlp(ts(newcases_us_multi$positiveIncrease[1:165]), hd = 5, comb = "median", xreg = newvar_fore_us[1:165,])
forecasts_us_mlp = forecast(mlp.fit,h = 12, xreg = newvar_fore_us[1:177,])

all_f = c(rep(1,4),forecasts_us_mlp$fitted, forecasts_us_mlp$mean)
plot(newcases_us_multi$positiveIncrease, type = "l")
lines(all_f, col = "blue")

ASE_final12 = mean((newcases_us_multi$positiveIncrease[166:177] -forecasts_us_mlp$mean)^2)
ASE_final12
#45799110


#Future Forecasts
mlp.fit = mlp(ts(newcases_us_multi$positiveIncrease), hd = 5, comb = "median", xreg = newvar_fore_us[1:177,])
short_us_mlp_m = forecast(mlp.fit,h = 7, xreg = newvar_fore_us[1:187,])
long_us_mlp_m = forecast(mlp.fit,h = 90, xreg = newvar_fore_us[1:267,])

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,187), main = "United States Short Term MLP Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(short_us_mlp_m$mean, col = "blue")

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,277), main = "United States Long Term MLP Multivariate Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(long_us_mlp_m$mean, col = "blue")




```


####US Ensemble

```{r}
ensemble_us_fore = (forecasts_us_mlp$mean + FinalPredictions_us_MLR)/2
plot(newcases_us_multi$positiveIncrease, type = "l")
lines(ensemble_us_fore, col = "blue")

#Final 12 ASE

ASE_final12 = mean((newcases_us_multi$positiveIncrease[166:177] -ensemble_us_fore)^2)
ASE_final12
#70596024

#Forecasting
#short_us_mlr_m  
#short_us_mlp_m

#long_us_mlr_m
#long_us_mlp_m

ensemble_us_fore_short = ( short_us_mlp_m$mean+ short_us_mlr_m)/2
ensemble_us_fore_long = (long_us_mlp_m$mean + long_us_mlr_m)/2

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,187), main = "United States Short Term Multivariate Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ensemble_us_fore_short, col = "blue")

plot(newcases_us$positiveIncrease, type = "l", xlim = c(1,277), main = "United States Long Term Multivariate Ensemble Forecasts", ylab = "Count of Cases", xlab= "Time")
lines(ensemble_us_fore_long, col = "blue")

```


##Conclusion


Given our data at hand we found our ARIMA(6,1,2) model to offer the best predictions for United States overall data. We found our Multiple Linear Regression model with correlated errors best predicted our Florida data. Given our current state of spread and recent trends of Covid19 spread, we expect increasing not optimistic forecasts of new case counts to show as most accurate in the coming days and months. 
